Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-02-14 numerical-analysis.tex
adding section on tracer particles

>>>>

One of the inherent drawbacks of a grid-based fluid method is the
inability to follow the evolution of a single parcel of fluid as it
travels through the simulation volume.  To address this, Enzo has the
capability to introduce Lagrangian ``tracer particles'' into a
calculation either at the beginning of the simulation or when
restarting the calculation.  These tracer particles are put into the
simulation in a rectangular solid volume with uniform, user-specified
spacing.  Each particle's position and velocity is updated over the
course of a single timestep $\Delta t$ as follows:

\begin{eqnarray}
\label{eqn.drifttrace}
x^{n+1/2} & = & x^n + (\Delta t/2) v^{interp,n} \nonumber \\
v^{n+1} & = & v^{interp,n+1} \\
x^{n+1} & = & x^{n+1/2} + (\Delta t/2) v^{n+1} \nonumber
\end{eqnarray}

This is essentially a drift-kick-drift particle update from time $n$
to time $n+1$ -- however, instead of computing an acceleration at the
half-timestep $t^n + \Delta t/2$ (as is done for massive particles --
see Equation~\ref{eqn.driftkick}), the particle's velocity is updated
both at the beginning of each timestep and at the half-timestep
by linearly interpolating the cell-centered baryon velocity to the
position of the particle ($v^{interp}$), and assigning it to that value.

At user-specified intervals, the tracer particles output their data
in a streaming format into a binary file.  The data written out
typically includes the tracer particle's unique ID, position, velocity, density,
and temperature, but is easily extensible to output any grid-based
quantity that a user requires.  This capability has been used quite
effectively in several papers, including \citep{2010ApJ...715.1575S, 2012ApJ...748...12S}.
